---
title: "PS07"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Michel Ruiz-Fuentes
SDS192 Data Science 
6 Dec. 2021

!["Figure 1. (2019). [Du Bois Legacy Committee Seeks Nominations for Second W.E.B. Du Bois Legacy Award] [Digital Art]. Great Barrington Massachussetts. https://www.townofgb.org/home/news/du-bois-legacy-committee-seeks-nominations-second-web-du-bois-legacy-award"](https://www.townofgb.org/sites/g/files/vyhlif636/f/styles/news_image_teaser/public/news/dubois-interior1-1540x1026_1.jpg?itok=75EiDw5S){width=85%}

## Inspired by W.E.B Du Bois' Legacy to Data Viz

Last year, I was enrolled in a statistics course named "Communicating with Data". I had an exciting experience learning how I can apply my **passion for civic engagement, creativity, and problem-solving** in the Statistics and Data Science department at Smith. We had several engaging units, but our focus on accessibility was one of my favorites. In this unit, we explored the importance of making data accessible for people of all backgrounds. For example, we discussed designing colorblind-friendly data visualizations by using inclusive color palettes.

This unit inspired me to search for more literature about this subject of **equity and accessibility in data visualizations**. I found the book "W.E.B. Du Bois' Data Portraits: Visualizing Black America" which includes map, graphs and charts that DuBois and a team of black sociologist used to illustrate the experiences of black Americans from 1790-1870. Ever since I discovered the data visualization contributions of DuBois and his team, I have been fascinated in using this tool to address agendas that relate to social good. In probleset 07, I will pursue this by engaging in the #DuBoisChallenge.  

## Where is the Data From

The data I am using, **W.E.B Du Bois Challenge** (published 02-16-2021), is from TidyTuesday. The challenge came from Anthony Starks in his article, [Recreating W.E.B Du Bois's Data Portraits](https://medium.com/nightingale/recreating-w-e-b-du-boiss-data-portraits-87dd36096f34). As he describes it, the goal of the #DuBoisChallenge is to "celebrate the data visualization legacy of W.E.B DuBois by recreating the visualizations from the 1900 Paris Exposition using modern tools."

## Loading necessary package and csv files from TidyTuesday

```{r}
#loading necessary packages
library(ggplot2)
library(knitr)
```

occupation.csv (Occupation by race)
```{r, show_col_types = FALSE}
occupation <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-02-16/occupation.csv')
```

city_rural.csv (Black population split between city and rural areas)
```{r, show_col_types = FALSE}
city_rural <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-02-16/city_rural.csv')
```


## Data Visualization 1: on City-Rural Residency 

```{r, echo = TRUE}
ggplot(data = city_rural, mapping = aes(x=Category, y= Population, fill= Category)) + 
  geom_col() + labs(title= "Where Did Black Americans Reside in the 1890s?",
  subtitle= "Population in City versus Rural Areas") + 
  xlab("Geographic Classification") + ylab("Population Amount") +
  theme(axis.text.x = element_text(angle = 20, hjust=1)) +
  scale_fill_manual(values = c("deepskyblue3","lightpink3", "seagreen3", "hotpink4")) 
```
## Viz 1 Debrief

This bar chart is a modern visualization of the data captured by Atlanta Univeristy and Du Bois' their visualization, [City and Rural Population in  1890](https://medium.com/nightingale/recreating-w-e-b-du-boiss-data-portraits-87dd36096f34). This visualization is one of several graphs that illustrates institutional racism and the racial disparities that Black Americans dealt with. How does this data convey inequities? By plotting geographic data we are highlighting issues assocaited with segregation and this consequently also exposes unequal opportunities Black Americans may have due to where they live. 

## Data Vizualization 2: on Occupations

```{r, echo = TRUE}
ggplot(data = occupation, mapping = aes(x=Percentage, y= Occupation, fill= Group)) + 
  geom_col() + labs(title= "Occupations Between Black and White Population",
  subtitle= "Data Collected in Georgia from 1790s to 1870s",
  fill= "Racial Groups") + 
  xlab("Percentage") + ylab("Occupation Type") +
  scale_fill_manual(values = c("salmon4", "bisque1"))
```


This bar chart demonstrates the differences between the types of jobs and occupations between the [Occupations of Black and WHite People in Georgia](https://www.loc.gov/item/2005676812/). This visualization is vital for our analysis of inequities because it shows how certain professions, especially ones with a reputation of paying better, were mostly for White Americans, whereas Black Americans filled the agricultural or domestic service roles. Our previous plot is instrumental to this analysis because we can assess that the majority of Black Americans lived in the "Country and Village" so it makes sense that the majority of their occupations were "Agriculture, Fisheries and Mining" and Domestic and Personal Service."

## Conclusions 

However, as Du Bois highlights in his studies, these data visualizations reveal social inequities and in particular illustrate the effects of institutional racism. At the end of this challenge, the challenge creator Anthony said this experience taught him to use data, "to tell compelling stories, and not to be afraid to shape your tools," and he is inspired by the [Du Bois illustrations](https://www.itsnicethat.com/articles/w-e-b-du-bois-charting-black-lives-house-of-illustration-131119) and hopes this creations can also inspire and motivate like Du Bois' legacy has. I seek to have the same impact in PS07 and other visualizations I create in the future. 

## Acknowledgements

I want to thank Professor Kim, Professor Poirier, and Professor Kinnaird for their support in building my data analysis skills!

